---
title: "ST512-SP26-Homework-3"
format: pdf
author: "Jason Menjivar"
---

## Problem 1

**a**

$\hat{y}$ = 25 - 0.5(7) = 21.5

**b**

$\hat{y}$ = 25 - 0.5(3) = 23.5

e = y - $\bar{y}$ = 30 - 23.5 = 6.5

The residual has a value of 6.5. Given that the residual is positive, the point will lie above the regression line. This is because our observed value of 30 is greater than the predicted value of 23.5.

**c**

In our model or fitted equation, the slope is -0.5. Thus, for every 1 unit increase in x, $\hat{y}$ will decrease by 0.5. If x is increasing by 3 units, this would then mean that $\hat{y}$ will decrease by 1.5 units.

-0.5(3) = -1.5

**d**

$\hat{y}$ = 25 - 0.5(6) = 22

Based on the fitted equation, the predicted value is 22. However, this is only a **predicted** value, not an actual observed value. There is random error associated with the model, thus the actual observed test score could differ from 22.

**e**

SSE = 7

n = 16

P = 2 (2 parameters, slope and intercept)

$\sigma^2$ = $\frac{SSE}{n-P}$ = $\frac{7}{14}$ = 0.5

## Problem 2

**a**

|            |     |         |        |        |          |
|------------|-----|---------|--------|--------|----------|
| Source     | DF  | SS      | MS     | F      | P        |
| Regression | 1   | 8654.7  | 8654.7 | 102.35 | 1.11e-15 |
| Error      | 75  | 6342.1  | 84.56  | --     | --       |
| Total      | 76  | 14996.8 | --     | --     | --       |

**b** 

$R^2$ = $\frac{SSR}{SST}$ = $\frac{8654.7}{6342.1}$ = 0.577

Approximately 57.7% of variation in rating can be  explained by the sugar content (in g).

**c** 

$\sigma^2$ = 84.56 

**d** 

Based on the calculated value in our ANOVA table, the p-value is very small <0.001. If we use a standard significance level of 0.05, we can reject the null hypothesis. Therefore, concluding that there is evidence that sugar content is a significant predictor of cereal ratings. 

## Problem 3 

**a**

t.value = $\frac{\hat{\beta_1}}{sd(\hat{\beta_1})}$ = $\frac{-1.867}{0.346}$ = -5.396 

p.value = 2 * p(t > |t.value|) = .00004

Using a standard significance level of 0.05, our estimated p.value is much smaller. Therefore, reject the null hypothesis, there is a significant relationship between treadmill time and race time. 

**b** 

CI: $\hat{\beta_1} \pm t_{1-\alpha/2, df_{error}} * sd(\hat{\beta_1})$ 

t.stat = 2.101

CI: -1.867 $\pm$ 2.101 * 0.346

95% CI: (-2.594, -1.140) 

**c**

58.816 -1.867(10) = 40.146 = point estimate 

CI: 40.146 $\pm$ 2.101 * 0.7342

CI: (38.58, 41.71)

**d** 

CI: 40.146 $\pm$ 2.101 * 2.226

CI: (35.47, 44.82) 

## Problem 4 



## R code 

**problem 1**

```{r}
1-pf(102.35, 1, 75)
```
**problem 3** 

```{r}
2*(1-pt(5.396,18))

qt(0.975, 18)
```

